Reproduce anaylses of Beninca et al (2008)

### 2016-02-23 code/report.Rmd
new method for calc. r-squared

@@

Load and run the functions, or read in from data file.

```{r}
# read script lines from website
script <- getURL(paste0(repo,"/code/indirect_method_functions.R"), ssl.verifypeer = FALSE)

# parase lines and evealuate in the global environement
eval(parse(text = script))

## To test the reproduction uncomment the next line of code
## don't run this unless you have some time (it takes a while)
#LE <- Get_GLE_Beninca(X)
#save(LE, file="~/Desktop/GLE_estimate.Rdata")

## load the already saved data from github (this can take some time depending on the internet connection)
source_data(paste0(repo, "/data/reproduction/GLE_estimate.Rdata?raw=True"))

LE[[1]]
```

This is quite far from the number using code and data from Steve (0.08415112). But recall that the functions have been carefully checked. Probably it is the data going into this function. A final step would be to save the data here, and take it into Steve's function. (This point is an issue on github.)

@@

```{r, echo=FALSE}
## put the models together into a list
all.gams <- LE[[2]]
#names(all.gams)
```

@@

And plot our version of figure 2:
```{r}
cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
                             "Nanophytoplankton", "Filamentous.diatoms", "Soluble.reactive.phosphorus",
                             "Total.dissolved.inorganic.nitrogen", "Bacteria", "Ostracods", "Harpacticoids",
                             "Protozoa"))

>>>>

The process is to transform the data, to fit GAMs, and then to calculated the global Lyapunov exponent.

@@

Load and run the functions, or read in from data file. Chunk not run by default, as take a minute or so. At last run, gave GLE of 0.03730891

```{r, eval=F}
# read script lines from website
script <- getURL(paste0(repo,"/code/indirect_method_functions.R"), ssl.verifypeer = FALSE)
eval(parse(text = script))

## Fit the GAMs
gams <- Fit_GAMs(X, gval=1.4, T_lag=1)

## calculate and print the GLE
Get_LE_from_fit2(gams, X1, epsval=0.01)

This is quite far from the number using code and data from Steve (0.08415112). This is caused by small differences in the data (and not by the functions), as we've found in using the code in the folder "GAM_test".

@@

The procedure is to fit GAMs (neural nets in the original publication) with response variable a function of explanatory variables at lag T, then to record r-squared of observed and fit data. Chunk not run by default, due to time required.

```{r, echo=FALSE, eval=FALSE}

T_lags <- 12
num_spp <- length(gams)
rsq_vals <- matrix(NA, T_lags, num_spp)
dimnames(rsq_vals)[[2]] <- names(gams)

for(i in 1:T_lags) {
  gams <- Fit_GAMs(X, gval=1.4, T_lag=i)
  for(j in 1:num_spp)
    rsq_vals[i,j] <- cor(gams[[j]]$fitted.values, gams[[j]]$model[1])^2
}

save(rsq_vals, file="~/Desktop/rsq_vals.Rdata")

<!--
Until the close HTML comment are some code Owen was previously using to calculate the predictability decline.

@@

And plot our version of figure 2:
```{r}
## Load the r-squared values that were previously saved (above)
## load the already saved data from github (this can take some time depending on the internet connection)
source_data(paste0(repo, "/data/reproduction/rsq_vals.Rdata?raw=True"))

cors <- data.frame(dist=1:12, rsq_vals)
cors.long <- gather(cors, key=dist)
names(cors.long) <- c("Prediction_distance", "Variable", "Correlation")

cors.long$Variable <- factor(cors.long$Variable, c("Cyclopoids", "Rotifers", "Calanoid.copepods", "Picophytoplankton",
                             "Nanophytoplankton", "Filamentous.diatoms", "Soluble.reactive.phosphorus",
                             "Total.dissolved.inorganic.nitrogen", "Bacteria", "Ostracods", "Harpacticoids",
                             "Protozoa"))
